---
title: "Bin Coordinate Data - Kmeans Clustering"
output: html_document
---

```{r setup, include=FALSE}
# LAST UPDATED: 2020/07/23
# Scott Carnie-Bronca
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Box/MPhil") #mac: "~/Box/MPhil" windows: "C:/Users/Scott/Box/MPhil/"

#Libraries
library(tidyverse)
library(lubridate)
library(sp)
library(leaflet)
library(gridExtra)

#Functions
source('~/Box/MPhil/Likelihood Estimation/bin_coords_kmeans.R')

#R options
options("lubridate.week.start"=1) #set first day of week to Monday instead of Sunday
```

## Load data

```{r loadData}
data = as_tibble(read.csv("Data Analysis/ridesDataAnalysis.csv", row.names = 1))
data

dataS = as_tibble(read.csv("Data Analysis/snapshotDataAnalysis.csv", row.names = 1))
dataS
```

Remove rows with invalid/unreasonable coordinates. Note: some rows appear to have coordinates swapped.
```{r removeInvalidCoords}
data = data %>%
  filter(pickupGeo0 >= 138.585 & pickupGeo0 <= 138.62 & pickupGeo1 >= -34.94 & pickupGeo1 <= -34.915 & +
        dropoffGeo0 >= 138.585 & dropoffGeo0 <= 138.62 & dropoffGeo1 >= -34.94 & dropoffGeo1  <= -34.915)
data
```

## Bin Data
Bin the data using the non-built in R function bin_coords_kmeans. Minimum/maximum coordinates are the tight CBD bounds. NOT WORKING, NEED TO FIX bin_coords_kmeans.R
```{r binData}
k = 5 #number of clusters
dataCoordsP = tibble(lat = data$pickupGeo1, lon = data$pickupGeo0, type = as.factor(rep("P", nrow(data))))
dataCoordsD = tibble(lat = data$dropoffGeo1, lon = data$dropoffGeo0, type = as.factor(rep("D", nrow(data))))
dataCoords = rbind(dataCoordsP, dataCoordsD)

# Cluster the data - see bin_coords_kmeans.R. Both pickups and dropoffs are used.

L = bin_coords_kmeans(select(dataCoords, lat, lon), k)
clusters = L[[1]]
L = L[[2]]
dataCoords = dataCoords %>%
  mutate(cluster = clusters$cluster)

pickupCluster = dataCoords %>%
  filter(type == "P") %>%
  select(cluster)
dropoffCluster = dataCoords %>%
  filter(type == "D") %>%
  select(cluster)

#add the cluster indices to data
dataCell = data %>%
  mutate(pickupCluster = pickupCluster$cluster, 
         dropoffCluster = dropoffCluster$cluster)
```

## Plot clusters
```{r plotClusters}

#Each cluster
ggplot(dataCell, aes(x = pickupGeo0, y = pickupGeo1, color = factor(pickupCluster))) +
  geom_point(alpha = 0.2, size = 0.25) +
  geom_point(aes(x = dropoffGeo0, y = dropoffGeo1, color = factor(dropoffCluster)), alpha = 0.2, size = 0.25) +
  xlab("Longitude") + ylab("Latitude") + ggtitle("Each cluster")

#Where do the trips that end in each cluster come from
ggplot(dataCell, aes(x = dropoffGeo0, y = dropoffGeo1, color = factor(pickupCluster))) +
  geom_point(alpha = 0.2, size = 0.5) +
  xlab("Longitude") + ylab("Latitude") + ggtitle("Starting clusters")

#Where do the trips that start in each cluster go to
ggplot(dataCell, aes(x = pickupGeo0, y = pickupGeo1, color = factor(dropoffCluster))) +
  geom_point(alpha = 0.2, size = 0.5) +
  xlab("Longitude") + ylab("Latitude") + ggtitle("Ending clusters")
```


## Construct lists with each
A new data frame is constructed with pickup/dropoff times and the cell numbers for each trip.
```{r constructDataFrameByCell}
#Sort by time in each cell
dataCell = dataCell %>%
  group_by(pickupCluster, dropoffCluster) %>%
  arrange(pickupTs)

dataCell
```

## Plot movement between cells
Plot the inter-cell movement:
```{r plotMovementBetweenCells}
#nInCell = summarise(dataCell, n = length(pickupTime))
ggplot(dataCell, aes(x = pickupCluster, y = dropoffCluster)) +
  #scale_fill_distiller(palette= "Set3", direction=1) +
  geom_bin2d(binwidth = 1)
```

The diagonal line has high counts, indicating that a high proportion of scooter trips stay within each cell.

## Select a cell and generate list of hires/dropoffs
Generate list of events in a specific cell.
```{r individualCell}
cellNum = 1 #cell number, from 1 to k

#Pickup events:
pickupTimes = dataCell %>%
  group_by() %>%
  filter(pickupCluster == cellNum) %>%
  select(pickupTs)
pickupTimes = ymd_hms(pickupTimes$pickupTs)

dropoffTimes = dataCell %>%
  group_by() %>%
  filter(dropoffCluster == cellNum) %>%
  select(dropoffTs)
dropoffTimes = ymd_hms(dropoffTimes$dropoffTs)

eventTimes = c(pickupTimes, dropoffTimes)
eventTypes = c(rep("H", length(pickupTimes)), rep("D", length(dropoffTimes)))
events = tibble(times = eventTimes, types = as.factor(eventTypes))
events = events%>%
  arrange(times)
events
```

## Bin snapshot data
The snapshot data is binned using the clusters from the rides data:
```{r binSnapshots}
dataS = as_tibble(read.csv("Data Analysis/snapshotDataAnalysis.csv", row.names = 1))
coordsS = dataS %>%
  select(geo0, geo1) %>%
  coordinates()

centroids = L[[2]]
centroids[,1:2] = centroids[,2:1] #swap columns; should be lon then lat

centroidDistance = spDists(coordsS, centroids, longlat = TRUE) #in km

snapshotCluster = rep(0, nrow(centroidDistance))
for(k in seq(1:nrow(centroidDistance))){
  snapshotCluster[k] = which.min(centroidDistance[k,])
}

dataS = dataS %>%
  mutate(cell = snapshotCluster)
```

## Save binned data

```{r saveData}
write.csv(dataCell, "Likelihood Estimation/binnedData_kmeans.csv")
write.csv(dataS, "Likelihood Estimation/binnedSnapshotData_kmeans.csv")
```
