---
title: "Bin Coordinate Data - Beam"
output: html_document
---

```{r setup, include=FALSE}
# LAST UPDATED: 2021/03/29
# Scott Carnie-Bronca
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "../") #mac: "~/Box/MPhil" windows: "C:/Users/Scott/Box/MPhil/"

#Libraries
pacman::p_load(tidyverse, lubridate, sp, leaflet)
#Functions

source('../Likelihood Estimation/bin_coords.R')

#R options
options("lubridate.week.start"=1) #set first day of week to Monday instead of Sunday
options(pillar.sigfig = 7)
```

## Load data

```{r loadData_rides}
data = as_tibble(read.csv("../Data Analysis/beam_ridesCleaned.csv", row.names = 1))
data
```

## Bin Data
Bin the data using the non-built in R function bin_coords. Minimum/maximum coordinates are the tight CBD bounds. 
```{r binData_rides}
n = 50 #n bins on each axis
minlat = -34.9425
maxlat = -34.8975
minlon = 138.575
maxlon = 138.625

# Get distance in km between points
lat = rep(c(minlat, maxlat), 2)
lon = rep(c(minlon, maxlon), each = 2)
X = cbind(lon, lat)
dist = spDists(X, longlat = TRUE)
dist
X %>% leaflet() %>% addTiles() %>% addMarkers()

# Filter to binning range
data = data %>%
  filter(start_lat > minlat, start_lat < maxlat, start_lon > minlon, start_lon < maxlon,
         end_lat > minlat, end_lat < maxlat, end_lon > minlon, end_lon < maxlon)


pickupLat = data$start_lat
pickupLon = data$start_lon

pickupList = bin_coords(pickupLat, pickupLon, n, minlat, maxlat, minlon, maxlon)
pickupInd = pickupList[[1]]
pickupBreaks = pickupList[[2]]
pickupInd
pickupBreaks


dropoffLat = data$end_lat
dropoffLon = data$end_lon

dropoffList = bin_coords(dropoffLat, dropoffLon, n, minlat, maxlat, minlon, maxlon)
dropoffInd = dropoffList[[1]]
dropoffBreaks = dropoffList[[2]]
dropoffInd
dropoffBreaks

ggplot() +
  geom_bin2d(aes(x = longitudeInd, y = latitudeInd), rbind(pickupInd, dropoffInd), binwidth = 1)
```

## Construct lists with each
A new data frame is constructed with pickup/dropoff times and the cell numbers for each trip.
```{r constructDataFrameByCell_rides}
#Construct data frame
dataCell = data %>%
  add_column(start_lat_ind = pickupInd$latitudeInd,
             start_lon_ind = pickupInd$longitudeInd,
             end_lat_ind = dropoffInd$latitudeInd,
             end_lon_ind = dropoffInd$longitudeInd)
                 
#Add cell number
dataCell = dataCell %>%
  mutate(pickup_cell_num = (start_lat_ind-1) + n*(start_lon_ind-1),
         dropoff_cell_num = (end_lat_ind-1) + n*(end_lon_ind-1))
#Sort by time in each cell
dataCell = dataCell %>%
  group_by(pickup_cell_num, dropoff_cell_num) %>%
  arrange(ymd_hms(start_time))

dataCell
```

## Plot movement between cells
The inter-cell movement is plotted. 
```{r plotMovementBetweenCells_rides}
#nInCell = summarise(dataCell, n = length(pickupTime))
ggplot(dataCell, aes(x = pickup_cell_num, y = dropoff_cell_num)) +
  scale_fill_distiller(palette= "Set3", direction=1) +
  geom_bin2d(binwidth = 1)

dataCell %>%
  group_by(pickup_cell_num, dropoff_cell_num) %>%
  summarise(n = n(), start_lat = mean(start_lat), start_lon = mean(start_lon), end_lat = mean(end_lat), end_lon = mean(end_lon)) %>% 
  arrange(-n)


top_trips = dataCell %>%
  group_by(pickup_cell_num, dropoff_cell_num) %>%
  summarise(n = n(), start_lat = mean(start_lat), start_lon = mean(start_lon), end_lat = mean(end_lat), end_lon = mean(end_lon)) %>% 
  ungroup() %>%
  arrange(-n) %>%
  slice_max(n, n = 10) %>%
  transmute(trip = paste(pickup_cell_num, dropoff_cell_num))

dataCell %>% 
  mutate(trip = paste(pickup_cell_num, dropoff_cell_num)) %>%
  filter(trip %in% top_trips$trip) %>%
  ggplot(aes(x = trip_duration, col = trip)) + geom_density(alpha = 0.5) +
  geom_density(aes(x = trip_duration, col = NULL), data = dataCell)
```

## Look at tour vs. non-tour
```{r}
dataCell %>%
  mutate(is_tour = pickup_cell_num == dropoff_cell_num) %>%
  ggplot(aes(x = trip_duration, col = is_tour)) + geom_histogram(binwidth = 60)

dataCell %>%
  mutate(is_tour = pickup_cell_num == dropoff_cell_num) %>%
  ggplot(aes(x = trip_duration, y = trip_distance, col = is_tour)) + geom_point()

dataCell %>%
  mutate(is_tour = pickup_cell_num == dropoff_cell_num) %>%
  ggplot(aes(x = trip_distance, col = is_tour)) + geom_histogram(binwidth = 20)
```

## Save binned data

```{r saveData_rides}
write.csv(dataCell, "../Likelihood Estimation/trips_binned_beam.csv")
```


# Status data
## Load data

```{r loadData}
data = as_tibble(read.csv("../Data Analysis/beam_statusCleaned.csv", row.names = 1))
data
```

## Bin Data
Bin the data using the non-built in R function bin_coords. Minimum/maximum coordinates are the tight CBD bounds. 
```{r binData}
# Filter to binning range
data = data %>%
  filter(lat > minlat, lat < maxlat, lon > minlon, lon < maxlon)


lat = data$lat
lon = data$lon

binnedList = bin_coords(lat, lon, n, minlat, maxlat, minlon, maxlon)
binnedInd = binnedList[[1]]
binnedBreaks = binnedList[[2]]

ggplot() +
  geom_bin2d(aes(x = longitudeInd, y = latitudeInd), binnedInd, binwidth = 1)
```

## Construct lists with each
A new data frame is constructed with pickup/dropoff times and the cell numbers for each trip.
```{r constructDataFrameByCell}
#Construct data frame
dataCell = data %>%
  add_column(lat_ind = binnedInd$latitudeInd,
             lon_ind = binnedInd$longitudeInd)
                 
#Add cell number
dataCell = dataCell %>%
  mutate(cell_num = (lat_ind-1) + n*(lon_ind-1))
#Sort by time in each cell
dataCell = dataCell %>%
  group_by(cell_num) %>%
  arrange(ymd_hms(event_time))

dataCell
```


## Save binned data

```{r saveData}
write.csv(dataCell, "../Likelihood Estimation/status_binned_beam.csv")
```
