---
title: "Route binning + speed"
output: html_document
---


# Status data
## Load data

```{r setup, include=FALSE}
# LAST UPDATED: 2021/05/19
# Scott Carnie-Bronca
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "../") #mac: "~/Box/MPhil" windows: "C:/Users/Scott/Box/MPhil/"

#Libraries
pacman::p_load(tidyverse, lubridate, sp, leaflet)
#Functions

source('../Likelihood Estimation/bin_coords.R')

#R options
options("lubridate.week.start"=1) #set first day of week to Monday instead of Sunday
options(pillar.sigfig = 7)
```

```{r loadData}
data = as_tibble(read.csv("../Data Analysis/beam_routesCleaned_trimmed.csv", row.names = 1))
data
```

## Bin Data
Bin the data using the non-built in R function bin_coords. Minimum/maximum coordinates are the tight CBD bounds. 
```{r binData}
n = 50 #n bins on each axis
minlat = -34.9425
maxlat = -34.8975
minlon = 138.575
maxlon = 138.625

# Filter to binning range
data = data %>%
  filter(lat > minlat, lat < maxlat, lon > minlon, lon < maxlon)


lat = data$lat
lon = data$lon

binnedList = bin_coords(lat, lon, n, minlat, maxlat, minlon, maxlon)
binnedInd = binnedList[[1]]
binnedBreaks = binnedList[[2]]

ggplot() +
  geom_bin2d(aes(x = longitudeInd, y = latitudeInd), binnedInd, binwidth = 1)
```

## Construct lists with each
A new data frame is constructed with pickup/dropoff times and the cell numbers for each trip.
```{r constructDataFrameByCell}
#Construct data frame
dataCell = data %>%
  add_column(lat_ind = binnedInd$latitudeInd,
             lon_ind = binnedInd$longitudeInd)
                 
#Add cell number
dataCell = dataCell %>%
  mutate(cell_num = (lat_ind-1) + n*(lon_ind-1),
         cell_num = (lat_ind-1) + n*(lon_ind-1))
#Sort by time in each cell
dataCell = dataCell %>%
  group_by(cell_num) %>%
  arrange(time)

dataCell
```

## Save binned data

```{r saveData}
write.csv(dataCell, "../Likelihood Estimation/routes_binned_beam.csv")
```

## Instantaneous Speed
```{r}
pacman::p_load(geodist)

# Add time and distance, then speed
dataCell = dataCell %>% ungroup()
dataCell = dataCell %>%
  group_by(trip_id) %>%
  mutate(t_diff = time-lag(time),
         dist = geodist_vec(lon, lat, sequential = TRUE, pad = TRUE),
         speed = 1000*dist/t_diff) %>%
  drop_na() %>%
  filter(speed < 20) #remove very high speed values



dataCell %>% ggplot(aes(x = time, y = speed)) + geom_point()

dataCell %>% ggplot(aes(x = speed)) + geom_density()

dataCell %>% ungroup() %>% mutate(day_of_week = factor(wday(as.POSIXct(time/1000, origin = origin))), is_weekend = (day_of_week == 6 | day_of_week == 7)) %>% ggplot(aes(x = speed, col = day_of_week)) + geom_density()

dataCell %>% ungroup() %>% slice_sample(n = 1e6) %>% mutate(day_of_week = factor(wday(as.POSIXct(time/1000, origin = origin))), is_weekend = (day_of_week == 6 | day_of_week == 7), time = factor(floor_date(as.POSIXct(time/1000, origin = origin), unit = "days"))) %>% ggplot(aes(y = speed, x = time, col = is_weekend)) + geom_violin()

dataCell %>% group_by(lon_ind, lat_ind) %>% summarise(speed = speed) %>% ggplot(aes(x = lon_ind, y = lat_ind, fill = speed)) + geom_raster()
```

Majority of trips are <5m/s, which agrees with the speed limit of about 15km/hr \approx 4.2m/s

