---
title: "Bin Coordinate Data"
output: html_document
---

```{r setup, include=FALSE}
# LAST UPDATED: 2021/10/28 - Updated to match grid size to Beam binning
# Scott Carnie-Bronca
knitr::opts_chunk$set(echo = TRUE)

#Libraries
library(tidyverse)
library(lubridate)
library(sp)
library(leaflet)
library(gridExtra)
library(skimr)

#Functions

source('../Likelihood Estimation/bin_coords.R')

#R options
options("lubridate.week.start"=1) #set first day of week to Monday instead of Sunday
```

## Load data

```{r loadData}
data = as_tibble(read.csv("../Data Analysis/ridesDataAnalysis.csv", row.names = 1))
data
```

Remove rows with invalid/unreasonable coordinates. Note: some rows appear to have coordinates swapped.
```{r removeInvalidCoords}
minlat = -34.9425
maxlat = -34.8975
minlon = 138.575
maxlon = 138.625

data = data %>%
  filter(pickupGeo0 >= minlon & pickupGeo0 <= maxlon & pickupGeo1 >= minlat & pickupGeo1 <= maxlat & +
           dropoffGeo0 >= minlon & dropoffGeo0 <= maxlon & dropoffGeo1 >= minlat & dropoffGeo1  <= maxlat)
data
```

## Bin Data
Bin the data using the non-built in R function bin_coords. Minimum/maximum coordinates are the tight CBD bounds. 
```{r binData}
n = 100 #n bins on each axis


pickupLat = data$pickupGeo1
pickupLon = data$pickupGeo0

pickupList = bin_coords(pickupLat, pickupLon, n, minlat, maxlat, minlon, maxlon)
pickupInd = pickupList[[1]]
pickupBreaks = pickupList[[2]]
pickupInd
pickupBreaks


dropoffLat = data$dropoffGeo1
dropoffLon = data$dropoffGeo0

dropoffList = bin_coords(dropoffLat, dropoffLon, n, minlat, maxlat, minlon, maxlon)
dropoffInd = dropoffList[[1]]
dropoffBreaks = dropoffList[[2]]
dropoffInd
dropoffBreaks

ggplot() +
  geom_bin2d(aes(x = longitudeInd, y = latitudeInd), rbind(pickupInd, dropoffInd), binwidth = 1)
```

## Construct lists with each
A new data frame is constructed with pickup/dropoff times and the cell numbers for each trip.
```{r constructDataFrameByCell}
#Construct data frame
dataCell = data %>%
  add_column(pickupLatInd = pickupInd$latitudeInd,
             pickupLonInd = pickupInd$longitudeInd,
             dropoffLatInd = dropoffInd$latitudeInd,
             dropoffLonInd = dropoffInd$longitudeInd) %>%
  rename(pickupTime = pickupTs,
         dropoffTime = dropoffTs)

#Add cell number
dataCell = dataCell %>%
  mutate(pickupCellNum = (pickupLatInd-1) + n*(pickupLonInd-1),
         dropoffCellNum = (dropoffLatInd-1) + n*(dropoffLonInd-1))
#Sort by time in each cell
dataCell = dataCell %>%
  group_by(pickupCellNum, dropoffCellNum) %>%
  arrange(pickupTime)

dataCell
```

## Plot movement between cells
The inter-cell movement is plotted. Note that nBins=n < n^2, so consecutive cells are plotted at the same point.
```{r plotMovementBetweenCells}
#nInCell = summarise(dataCell, n = length(pickupTime))
ggplot(dataCell, aes(x = pickupCellNum, y = dropoffCellNum)) +
  scale_fill_distiller(palette= "Set3", direction=1) +
  geom_bin2d(binwidth = 1)
```


It appears that most scooters stay in their own cells, from the strong diagonal line.

<!-- ## Select a cell and generate list of hires/dropoffs -->


<!-- ```{r individualCell} -->
<!-- cellNum = 4546 #cell number, from 0 to n^2-1 -->

<!-- #Pickup events: -->
<!-- pickupTimes = dataCell %>% -->
<!--   group_by() %>% -->
<!--   filter(pickupCellNum == cellNum) %>% -->
<!--   select(pickupTime) -->
<!-- pickupTimes = ymd_hms(pickupTimes$pickupTime) -->

<!-- dropoffTimes = dataCell %>% -->
<!--   group_by() %>% -->
<!--   filter(dropoffCellNum == cellNum) %>% -->
<!--   select(dropoffTime) -->
<!-- dropoffTimes = ymd_hms(dropoffTimes$dropoffTime) -->

<!-- eventTimes = c(pickupTimes, dropoffTimes) -->
<!-- eventTypes = c(rep("H", length(pickupTimes)), rep("D", length(dropoffTimes))) -->
<!-- events = tibble(times = eventTimes, types = as.factor(eventTypes)) -->
<!-- events = events%>% -->
<!--   arrange(times) -->
<!-- events -->
<!-- ``` -->

## Save binned data

```{r saveData}
write.csv(dataCell, "../Likelihood Estimation/binnedData.csv")
```


## Snapshot Data
```{r}
data.snapshot = as_tibble(read.csv("../Data Analysis/snapshotsCleaned.csv"))
data.snapshot = data.snapshot %>%
  mutate(DataDate = ymd_hms(DataDate)) %>%
  select(DataDate, geo0, geo1, number)
```

Filter the snapshot data by time. Due to the lack of snapshot data for earlier times and the jump in scooter numbers on June 17th, June 18th onwards is chosen for parameter estimation.
```{r}
t0 = as.POSIXct("2019-06-18 00:00:00") #initial time
tF = as.POSIXct("2019-7-10 03:00:00") #final time, greater than the last time in the snapshot data
data.snapshot = data.snapshot %>%
  filter(DataDate >= t0 & DataDate <= tF)
```

Filter the data by location - all points should be in Adelaide.
```{r}
data.snapshot = data.snapshot %>%
  filter(geo0 >= minlon & geo0 <= maxlon & geo1 >= minlat & geo1 <= maxlat)
data.snapshot
```

Bin the data, similar to the ride data
```{r}
lat = data.snapshot$geo1
lon = data.snapshot$geo0

list = bin_coords(lat, lon, n, minlat, maxlat, minlon, maxlon)
data.snapshot.binned = list[[1]]
breaks = list[[2]]
data.snapshot.binned
breaks

data.snapshot.binned = data.snapshot.binned %>%
  mutate(cellNum = latitudeInd-1 + n*(longitudeInd-1), DataDate = data.snapshot$DataDate, number = data.snapshot$number)
data.snapshot.binned
```

```{r}
write.csv(data.snapshot.binned, "../Likelihood Estimation/snapshotBinned.csv")
```

Plot of cells with scooters:
```{r}
data.snapshot.binned %>%
  group_by(DataDate, cellNum) %>%
  summarise(n_scooters = n()) %>%
  ungroup() %>%
  group_by(DataDate) %>%
  complete(cellNum = 0:(n^2-1), fill = list("n_scooters" = 0)) %>%
  summarise(p = sum(n_scooters > 0)/n^2) %>%
  ggplot(aes(x = DataDate, y = p)) + geom_line() + coord_cartesian(ylim = c(0, 0.25)) + labs(x = "Time", y = "Proportion of cells with scooters") + theme_bw()

ggsave("./cells_with_scooters_ride.pdf", width = 6, height = 4) 
```

# Sizes of cells
```{r}
x = matrix(c(minlon, maxlon, minlon, maxlon, minlat, minlat, maxlat, maxlat), ncol = 2)
spDists(x, longlat = TRUE)*(1000/n)
```

# Plot of corners
```{r}
pacman::p_load(mapview, sf)

#x %>% leaflet() %>% addTiles %>% addCircleMarkers() %>% mapshot(file = "./corners.png")
tibble(longitude = x[,1], latitude = x[,2]) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>%
  mapview(legend = FALSE)

tibble(longitude = x[,1], latitude = x[,2]) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>%
  mapview(legend = FALSE) %>%
  mapshot(file = "./corners.png")
```

```{r}
tibble(longitude = x[,1], latitude = x[,2]) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>%
  st_make_grid(n = c(4, 4)) %>%
  mapview(legend = TRUE)

tibble(longitude = x[,1], latitude = x[,2]) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>%
  st_make_grid(n = c(4, 4)) %>%
  mapview(legend = TRUE) %>%
  mapshot(file = "./smallgrid.png")

tibble(longitude = x[,1], latitude = x[,2]) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0") %>%
  st_make_grid(n = c(100, 100), square = FALSE) %>%
  mapview(alpha = 0.1)
```

# Save grid for plotting as spatialpolygons
```{r}
grid_corners = pickupBreaks %>% as.matrix #n+1 rows, col 1 = lat, col 2 = lon
polygon_list = list(n^2)
for(k in 1:n^2){
  lat_ind = ((k-1) %% n) + 1
  lon_ind = ((k-1) %/% n) + 1
  polygon_list[[k]] = matrix(
    c(
     grid_corners[lon_ind, 2],   grid_corners[lat_ind, 1],
     grid_corners[lon_ind, 2],   grid_corners[lat_ind+1, 1],
     grid_corners[lon_ind+1, 2], grid_corners[lat_ind+1, 1],
     grid_corners[lon_ind+1, 2], grid_corners[lat_ind, 1],
     grid_corners[lon_ind, 2],   grid_corners[lat_ind, 1]
    ),
    nrow = 5, byrow = TRUE
  ) %>%
    list() %>%
    Polygon() %>%
    list() %>%
    Polygons(ID = k-1)
}

spatial_polygons_list = polygon_list %>%
  SpatialPolygons(proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")) %>%
  st_as_sfc

spatial_polygons_list %>% mapView()

spatial_polygons_list %>% st_write("./grid.shp", append = FALSE)
```
