---
title: "Estimate demand - Beam data"
output: html_document
---

```{r setup, include=FALSE}
# LAST UPDATED: 2021/08/02
# Scott Carnie-Bronca
knitr::opts_chunk$set(echo = TRUE)

#Libraries
pacman::p_load(tidyverse, lubridate, sp, leaflet)

#R options
options("lubridate.week.start"=1) #set first day of week to Monday instead of Sunday
```


## Hire threshold: how many hires are required to get an estimate of demand. Time threshold: minimum time in seconds to get an estimate of demand
```{r}
pickup_threshold = 10
time_threshold = 3600
```

# Beam
## Load data
```{r}
n = 100 #number of bins on each axis

trips = read.csv("./trips_binned_beam.csv", row.names = 1) %>% as_tibble()
status = read.csv("./status_binned_r_beam.csv", row.names = 1) %>% as_tibble() %>% mutate(event_time = ymd_hms(event_time))
```

## Get the last locations for each scooter
```{r}
last_locations = status %>%
  arrange(location_timestamp) %>%
  group_by(vehicle_id) %>%
  slice_tail(n = 1) %>%
  ungroup()

last_locations %>% ggplot(aes(x = lon, y = lat)) + geom_point()
last_locations %>% ggplot(aes(x = event_time, col = event_type_reason)) + geom_histogram(binwidth = 3600*12)
last_populations = last_locations$cell_num %>% table()
last_populations
```

Using snapshot data:
```{r, eval = FALSE}
snapshot_filenames = list.files("../Data/Output/Snapshots/")
max_file_ind = min(which(ymd_hms(gsub(".json", "", snapshot_filenames), tz = "Australia/Adelaide") >= max(ymd_hms(status$event_time, tz = "Australia/Adelaide"), na.rm = TRUE))) # find the first snapshot that takes place after the status data

last_snapshot = jsonlite::read_json(paste0("../Data/Output/Snapshots/", snapshot_filenames[max_file_ind]), simplifyVector = TRUE) %>% as_tibble()

source("bin_coords.R")
minlat = -34.9425
maxlat = -34.8975
minlon = 138.575
maxlon = 138.625

get_cell_num = function(lat, lon) {
  X = bin_coords(lat, lon, n, minlat, maxlat, minlon, maxlon)
  out = (X[[1]]$latitudeInd-1) + n*(X[[1]]$longitudeInd-1)
  return(out)
}

last_snapshot = last_snapshot %>% 
  filter(lat > minlat, lat < maxlat, lon > minlon, lon < maxlon) %>%
  mutate(cell_num = get_cell_num(lat, lon))

last_snapshot %>% ggplot(aes(x = lon, y = lat)) + geom_point()
last_populations = last_snapshot$cell_num %>% table()
last_populations
```


## For each cell, find periods of time when scooters are not present.
```{r}
current_population = rep(0, n^2)
current_population[as.numeric(names(last_populations))+1] = last_populations
cell_status = tibble(cell = 0, n_pickups = 0, nzi_start = 0, nzi_end = 0)
nzi_end = rep(max(status$location_timestamp), n^2) #end time of non-zero-population interval
n_pickups = rep(0, n^2) #number of pickups since the previous time no scooters were available

status = status %>%
  mutate(change = case_when(
    event_type_reason %in% c("user_drop_off", "maintenance_drop_off", "redistribution_in", "extra_drop_off") ~ 1,
    event_type_reason %in% c("user_pick_up",  "maintenance_pick_up", "redistribution_out", "extra_pick_up") ~ -1,
    TRUE                                                                                   ~ 0
  ))

status %>%
  group_by(cell_num) %>%
  summarise(final_change = sum(change)) %>%
  pull(final_change) %>%
  table

status = status %>%
  filter(cell_num >= 0)

cell_populations = tibble(cell = 0, n_in_cell = 0, t = 0) #check scooter counts

for(k in nrow(status):1){
  t = status$location_timestamp[k] #what time did the event occur at
  cell = status$cell_num[k] #what cell did the event occur in
  change = status$change[k] #how did the scooter count change at this time
  type = status$event_type_reason[k]
  cell_population = current_population[cell+1] #cell population after the event
  current_population[cell+1] = current_population[cell+1] - change
  
  # Debugging
  cell_populations = cell_populations %>% add_row(cell = cell, n_in_cell = cell_population, t = t)
  #if(current_population[cell] < 0){print(paste(k, current_population[cell], cell, status$vehicle_id[k], type))}
  
  if(cell_population == 0){ 
    #print(1)
    #if the number of scooters reaches 0 after the event
   
    if(type == "user_pick_up"){n_pickups[cell+1] = n_pickups[cell+1] + 1} 
          
    nzi_end[cell+1] = t
    #print(paste("update nzi_end: ", cell, t, nzi_end[cell]))

    
  } else {
    #print(2)
    if(current_population[cell+1] == 0){
      #print(3)
      #if the cell population leaves 0 at the event

      # Update tibble
      #print(paste("update cell status: ", cell, n_pickups[cell], t, nzi_end[cell]))
      cell_status = cell_status %>%
        add_row(cell, n_pickups = n_pickups[cell+1], nzi_start = t, nzi_end = nzi_end[cell+1])
      n_pickups[cell+1] = 0
      nzi_end[cell+1] = 0
    } else {
      #print(4)
      #if the cell population is > 0 either side of the event
      if(type == "user_pick_up"){n_pickups[cell+1] = n_pickups[cell+1] + 1}
    }
  }
}
cell_status = cell_status[-1,]
cell_status
```

## Plot
```{r}
cell_populations = cell_populations %>% filter(cell > 0) %>% group_by(cell) %>% mutate(i = row_number())

cell_populations %>% ungroup() %>% mutate(i_overall = row_number()) %>% ggplot(aes(x = i_overall, y = n_in_cell, group = cell)) + geom_line()
cell_populations %>% ungroup() %>% mutate(i_overall = row_number()) %>% ggplot(aes(x = i, y = n_in_cell, group = cell)) + geom_line()
```

## Proportion of cells with scooters over time
```{r}
first_day = min(floor_date(as.POSIXct(status$location_timestamp/1000, tz = "Australia/Adelaide", origin = origin), "day"))
tz(first_day) <- "Australia/Adelaide"
last_day = max(floor_date(as.POSIXct(status$location_timestamp/1000, tz = "Australia/Adelaide", origin = origin), "day"))
tz(last_day) <- "Australia/Adelaide"
timeBreaks = seq(first_day, last_day, by = "DSTday")
timeBreaks = as.numeric(rep(timeBreaks, each = 48) + rep(1800*(0:47), times = length(timeBreaks)))*1000

cell_populations$t_regular = NA

cell_populations = cell_populations %>%
  ungroup() %>%
  add_row(t = timeBreaks, t_regular = timeBreaks) %>%
  arrange(t) %>%
  fill(t_regular, .direction = "up") %>%
  drop_na(cell)

cell_populations %>%
  group_by(t_regular, cell) %>%
  slice_tail() %>%
  group_by(t_regular) %>%
  complete(cell = 0:(n^2-1), fill = list(n_in_cell = 0)) %>%
  summarise(p_greater_zero = sum(n_in_cell > 0)/n^2, p_equal_zero = sum(n_in_cell == 0)/n^2) %>%
  mutate(t_regular = as.POSIXct(t_regular/1000, tz = "Australia/Adelaide", origin = origin), p_greater_zero = p_greater_zero*4) %>%
  ggplot(aes(x = t_regular, y = p_greater_zero)) + geom_line() +# + geom_line(aes(y = p_equal_zero), col = "red")
  labs(y = "Proportion of cells with scooters", x = "Time") + coord_cartesian(ylim = c(0, 0.25)) + theme_bw()
ggsave("./cells_with_scooters.pdf", width = 6, height = 4)
```

## Estimate demand
```{r}
cell_status = cell_status %>%
  mutate(demand = 1000*n_pickups/(nzi_end-nzi_start))

cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, unweighted_demand =mean(demand), weighted_demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  arrange(-unweighted_demand)
cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, unweighted_demand = mean(demand), weighted_demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  arrange(-weighted_demand)

cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, unweighted_demand =mean(demand), weighted_demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  ggplot(aes(x = unweighted_demand, y = weighted_demand, col = n_hires)) + geom_point() + geom_abline()

cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, unweighted_demand =mean(demand), weighted_demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  ggplot(aes(x = unweighted_demand, y = weighted_demand, col = log10(nonzero_time))) + geom_point() + geom_abline()

cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, unweighted_demand =mean(demand), weighted_demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  pivot_longer(cols = c("unweighted_demand", "weighted_demand"), names_to = "mean_type") %>%
  ggplot(aes(x = n_hires, y = value, col = mean_type)) + geom_point()
cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, unweighted_demand =mean(demand), weighted_demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  pivot_longer(cols = c("unweighted_demand", "weighted_demand"), names_to = "mean_type") %>%
  ggplot(aes(x = nonzero_time, y = value, col = mean_type)) + geom_point()
```

## Plot
```{r}
cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = mean(demand)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold, demand < 0.5) %>%
  complete(cell = 0:(n^2-1), fill = list(demand = 0, nonzero_time = 0, n_hires = 0)) %>%
  mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + geom_tile()

cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = n_hires/nonzero_time) %>%
  drop_na() %>%
  mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  #filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + geom_tile()

cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  #filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  mutate(lat_ind = ((cell + 1) %% n), lon_ind = ((cell) %/% n) + 1,
         #demand = case_when(
        #   demand > 0.0005 ~ 0.0005,
        #   TRUE           ~ demand
         #)
        ) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + geom_tile()
```

## Filter out inaccurate estimates
```{r}
cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = mean(demand)) %>%
  drop_na() %>%
  mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + geom_raster()



cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = n_hires/nonzero_time) %>%
  drop_na() %>%
  mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + geom_raster()

cell_status %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  mutate(lat_ind = ((cell + 1) %% n), lon_ind = ((cell) %/% n) + 1,
         #demand = case_when(
        #   demand > 0.0005 ~ 0.0005,
        #   TRUE           ~ demand
         #)
        ) %>%
  complete(lat_ind = 1:n, lon_ind = 1:n, fill = list(demand = 0)) %>%
  mutate(demand = demand*3600) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + 
  geom_raster() + labs(fill = "Demand (hires/hr)", x = "", y = "") + 
  scale_fill_gradient2(low = "#173603", mid = "#6eff0d", high = "#ffffff", midpoint = 1, na.value = "#290e00") + 
  coord_fixed(ratio = 1) + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank()
  )

# Save for Stoneham
ggsave("./demand_est.png", width = 5, height = 4) 
```

## Save demands
```{r}
demand = cell_status %>%
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups, na.rm = TRUE), nonzero_time = sum(nzi_end-nzi_start, na.rm = TRUE)/1000, demand = n_hires/nonzero_time) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  complete(cell = 0:(n^2-1), fill = list(demand = 0))

demands = matrix(demand$demand, byrow = FALSE, ncol = n) #convert demand to a matrix (for compatibility with earlier demand estimation)

demands %>% write.table("../Likelihood Estimation/estimatedLambda_beam.csv", sep = ",", row.names = FALSE, col.names = FALSE)
```

# Split by time of day

## Define times to split by
```{r}
first_day = min(floor_date(status$event_time, "day"))
tz(first_day) <- "Australia/Adelaide"
last_day = max(floor_date(status$event_time, "day"))
tz(last_day) <- "Australia/Adelaide"
timeBreaks = seq(first_day, last_day, by = "DSTday")
timeBreaks = rep(timeBreaks, each = 5) + rep(c(0, 6*3600, 12*3600, 15*3600, 19*3600), times = length(timeBreaks))

status = status %>%
  add_row(tibble(
    location_timestamp = 1000*as.numeric(timeBreaks),
    cell_num = 0,
    event_type_reason = "time_break"
  )) %>%
  arrange(location_timestamp)
```

## For each time, find periods of time when scooters are not present.
```{r}
current_population = rep(0, n^2)
current_population[names(last_populations) %>% as.numeric] = last_populations
cell_status = tibble(cell = 0, n_pickups = 0, nzi_start = 0, nzi_end = 0, reason = "temp")
nzi_end = rep(max(status$location_timestamp), n^2) #end time of non-zero-population interval
n_pickups = rep(0, n^2) #number of pickups since the previous time no scooters were available
in_interval = rep(FALSE, n^2)
status = status %>%
  mutate(change = case_when(
    event_type_reason %in% c("user_drop_off", "maintenance_drop_off", "redistribution_in", "extra_drop_off") ~ 1,
    event_type_reason %in% c("user_pick_up",  "maintenance_pick_up", "redistribution_out", "extra_pick_up") ~ -1,
    TRUE                                                              ~ 0
  ))

cell_populations = tibble(cell = 0, n_in_cell = 0) #check scooter counts

for(k in nrow(status):1){ #(nrow(status)-10000)
  
  t = status$location_timestamp[k] #what time did the event occur at
  #print(paste("loop", k))
  # If the row is a time break
  if(status$event_type_reason[k] == "time_break"){ #if there is a time split
    #print("time break")
    #temp_tibble = tibble(cell = 0, n_pickups = 0, nzi_start = 0, nzi_end = 0)
    temp_cell = c()
    temp_n_pickups = c()
    temp_nzi_start = c()
    temp_nzi_end = c()
    for(cell in which(in_interval)){#for each cell
      #print(paste("time break, cell", cell))
      if(in_interval[cell+1]){ #if the interval is being split
        #print("needs updating")
        temp_cell = c(temp_cell, cell)
        temp_n_pickups = c(temp_n_pickups, n_pickups[cell+1])
        temp_nzi_start = c(temp_nzi_start, t)
        temp_nzi_end = c(temp_nzi_end, nzi_end[cell+1])
        
        # temp_tibble = rbind(temp_tibble, tibble_row(cell, n_pickups = n_pickups[cell], nzi_start = t, nzi_end = nzi_end[cell]))
        nzi_end[cell+1] = t
        n_pickups[cell+1] = 0
      }
    }
    cell_status = cell_status %>% add_row(cell = temp_cell, n_pickups = temp_n_pickups, nzi_start = temp_nzi_start, nzi_end = temp_nzi_end, reason = "time")
    next
  }
  cell = status$cell_num[k] #what cell did the event occur in
  change = status$change[k] #how did the scooter count change at this time
  type = status$event_type_reason[k]
  cell_population = current_population[cell+1] #cell population after the event
  current_population[cell+1] = current_population[cell+1] - change
  cell_populations = cell_populations %>% add_row(cell = cell, n_in_cell = cell_population)
  if(cell_population == 0){ 
    #print(1)
    #if the number of scooters reaches 0 after the event
   
    if(type == "user_pick_up"){n_pickups[cell+1] = n_pickups[cell+1] + 1} 
          
    in_interval[cell+1] = TRUE
    nzi_end[cell+1] = t
    #print(paste("update nzi_end: ", cell, t, nzi_end[cell]))

    
  } else {
    #print(2)
    cell_population_before = cell_population - change
    if(cell_population_before == 0){
      #print(3)
      #if the cell population leaves 0 at the event

      # Update tibble
      #print(paste("update cell status: ", cell, n_pickups[cell], t, nzi_end[cell]))
      cell_status = cell_status %>%
        add_row(cell, n_pickups = n_pickups[cell+1], nzi_start = t, nzi_end = nzi_end[cell+1], reason = "data")
      in_interval[cell+1] = FALSE
      n_pickups[cell+1] = 0
      nzi_end[cell+1] = 0
    } else {
      #print(4)
      #if the cell population is > 0 either side of the event
      if(type == "user_pick_up"){n_pickups[cell+1] = n_pickups[cell+1] + 1}
    }
  }
}
cell_status = cell_status[-1,]
cell_status
```

```{r}
cell_status = cell_status %>% 
  mutate(nzi_start_time = as.POSIXct(nzi_start/1000, origin = origin, tz = "Australia/Adelaide"), 
         hour = hour(nzi_start_time), 
         day = wday(nzi_start_time))
cell_status = cell_status %>% mutate(desc = paste(
    case_when(
        hour %in% 0:5 ~ "EarlyMorning",
        hour %in% 6:11 ~ "Morning",
        hour %in% 12:14 ~ "EarlyAfternoon",
        hour %in% 15:18 ~ "Afternoon",
        hour %in% 19:23 ~ "Evening",
        TRUE ~ "NA"
    ), case_when(
        day %in% c(6,7) ~ "Weekend",
        day %in% 1:5 ~ "Weekday",
        TRUE ~ "NA"
    ),
    sep = ""
))

cell_status %>%
  group_by(desc, cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = n_hires/nonzero_time) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + geom_raster() + facet_wrap(~desc)

cell_status %>%
  group_by(desc, cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = n_hires/nonzero_time) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  ggplot(aes(x = nonzero_time, y = demand)) + geom_point() + facet_wrap(~desc) + scale_y_log10()

cell_status %>%
  group_by(desc, cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = n_hires/nonzero_time) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1,
         time_length = case_when(
           str_detect(desc, "^EarlyMorning") ~ 6,
           str_detect(desc, "^Morning") ~ 6,
           str_detect(desc, "^EarlyAfternoon") ~ 3,
           str_detect(desc, "^Afternoon") ~ 4,
           str_detect(desc, "^Evening") ~ 5
         )) %>%
  filter(nonzero_time > (as.numeric(last_day-first_day) * (3600/5) *  time_length)) %>% #if 3600/x, must have scooters at least 1/xth of the time
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + geom_raster() + facet_wrap(~desc)
```

## Save demands
```{r}
timeDesc = expand_grid(time = 0:23, day = 1:7) %>%
  mutate(desc = paste(
    case_when(
      time %in% 0:5 ~ "EarlyMorning",
      time %in% 6:11 ~ "Morning",
      time %in% 12:14 ~ "EarlyAfternoon",
      time %in% 15:18 ~ "Afternoon",
      time %in% 19:23 ~ "Evening",
      TRUE ~ "NA"
    ), case_when(
      day %in% c(6,7) ~ "Weekend",
      day %in% 1:5 ~ "Weekday",
      TRUE ~ "NA"
    ),
    sep = ""
  ))

for(k in unique(timeDesc$desc)){
  demand_out = cell_status %>%
    filter(desc == k) %>%
    group_by(cell) %>%
    summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = n_hires/nonzero_time) %>%
    drop_na() %>%
    filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
    complete(cell = 1:n^2, fill = list(demand = 0))
  demand_out$demand %>% 
    matrix(byrow = FALSE, ncol = n) %>% 
    write.table(paste0("./Demands/demand_", k, "_beam.csv"), sep = ",", row.names = FALSE, col.names = FALSE)
}
```

# Plot cell populations
```{r}
cell_populations = cell_populations %>% filter(cell > 0) %>% group_by(cell) %>% mutate(i = row_number())

cell_populations %>% ggplot(aes(x = i, y = n_in_cell, group = cell)) + geom_line()

cell_populations_summary = cell_populations %>% 
  group_by(cell) %>%
  summarise(min_n = min(n_in_cell), max_n = max(n_in_cell), mean_n = mean(n_in_cell))

cell_populations_summary %>% mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = -min_n)) + geom_raster()
cell_populations_summary %>% mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = max_n)) + geom_raster()
cell_populations_summary %>% mutate(lat_ind = ((cell) %% n) + 1, lon_ind = ((cell) %/% n) + 1) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = mean_n)) + geom_raster()
```

