---
title: "Estimate demand - Beam data"
output: html_document
---

```{r setup, include=FALSE}
# LAST UPDATED: 2021/08/31
# Scott Carnie-Bronca
knitr::opts_chunk$set(echo = TRUE)

#Libraries
pacman::p_load(tidyverse, lubridate, sp, leaflet, foreach)

#R options
options("lubridate.week.start"=1) #set first day of week to Monday instead of Sunday
doParallel::registerDoParallel(cores = parallel::detectCores())
```


## Hire threshold: how many hires are required to get an estimate of demand. Time threshold: minimum time in seconds to get an estimate of demand
```{r}
pickup_threshold = 0
time_threshold = 0
```

# Ride
## Load data
```{r}
n = 50 #number of bins on each axis

trips = read.csv("./binnedData.csv", row.names = 1) %>% as_tibble() 
snapshots = read.csv("./snapshotBinned.csv", row.names = 1) %>% as_tibble()
```

## Reformat trip data
```{r}
trips = trips %>% 
  select(vehicleNumber, pickupTime, dropoffTime, pickupCellNum, dropoffCellNum) %>%
  pivot_longer(cols = c(pickupCellNum, dropoffCellNum)) %>%
  transmute(vehicle_id = vehicleNumber,
            event_type = str_extract(name, "pickup|dropoff"),
            event_time = case_when(
              event_type == "pickup" ~ pickupTime,
              event_type == "dropoff" ~ dropoffTime
            ),
            event_time_numeric = as.numeric(ymd_hms(event_time, tz = "Australia/Adelaide")),
            cell_num = value
  )
```

## Reformat snapshot data
```{r}
snapshots = snapshots %>%
  transmute(time = ymd_hms(DataDate, tz = "Australia/Adelaide"), cell_num = cellNum, times_numeric = as.numeric(time))
```

## Define times - snapshot times
```{r}
times = unique(snapshots$times_numeric)
n_times = length(times)-1
```

## Add columns to status data frame
```{r}
trips = trips %>%
  mutate(event_time = ymd_hms(event_time),
         change = case_when(
           event_type %in% c("dropoff") ~ 1,
           event_type %in% c("pickup")  ~ -1,
           TRUE                         ~ 0
         )) %>%
  arrange(event_time_numeric)
```

## Complete demand estimation code
```{r}
demands = foreach(j = 1:n_times, .combine = c, .packages = c("lubridate", "tidyverse")) %dopar% {
  cell_populations = tibble(cell = 0, n_in_cell = 0, t = 0, time_window = 0) #check scooter counts
  # Get INITIAL scooter locations
  t_start = times[j]
  t_end = times[j+1]
  
  # Find data points that are useful for finding the location of each scooter
  
  initial_locations = snapshots %>% filter(time == t_start) %>%
    pull(cell_num) %>%
    table
  
  # Loop through
  cells_with_scooters = 0:(n^2-1) %in% names(initial_locations) #which cells have scooters
  trips_month = trips %>% filter(event_time_numeric >= times[j], event_time_numeric <= times[j+1])
  next_population = rep(0, n^2)
  next_population[names(initial_locations) %>% as.numeric + 1] = initial_locations
  cell_status = tibble(cell = NA_real_, n_pickups = 0, nzi_start = 0, nzi_end = 0)
  nzi_start = rep(times[j], n^2)*cells_with_scooters #start time of non-zero-population interval
  n_pickups = rep(0, n^2) #number of pickups since the previous time scooters were available
  
  if(nrow(trips_month) > 0){ #if things happen
    for(k in 1:nrow(trips_month)){
      t = trips_month$event_time_numeric[k] #what time did the event occur at
      cell = trips_month$cell_num[k] #what cell did the event occur in
      change = trips_month$change[k] #how did the scooter count change at this time
      type = trips_month$event_type[k]
      cell_population = next_population[cell+1] #cell population before the event
      next_population[cell+1] = next_population[cell+1] + change #cell population after the change
      
      # Debugging
      cell_populations = cell_populations %>% add_row(cell = cell, n_in_cell = cell_population, t = t, time_window = j)
      
      
      if(cell_population == 0){
        #print(1)
        #if the number of scooters leaves 0 after the event
        
        #if(type == "user_pick_up"){n_pickups[cell+1] = n_pickups[cell+1] + 1}
        cells_with_scooters[cell+1] = TRUE
        nzi_start[cell+1] = t
        #print(paste("update nzi_start: ", cell, t, nzi_start[cell+1]))
        
        
      } else {
        #print(2)
        if(next_population[cell+1] == 0){
          #print(3)
          #if the cell population reaches 0 after the event
          
          if(type == "pickup"){n_pickups[cell+1] = n_pickups[cell+1] + 1}
          
          # Update tibble
          #print(paste("update cell status: ", cell, n_pickups[cell+1], t, nzi_start[cell+1]))
          cells_with_scooters[cell+1] = FALSE
          cell_status = cell_status %>%
            add_row(cell, n_pickups = n_pickups[cell+1], nzi_start = nzi_start[cell+1], nzi_end = t)
          n_pickups[cell+1] = 0
          nzi_start[cell+1] = -1 #this should be reset after the scooter count leaves 0
        } else {
          #print(4)
          #if the cell population is > 0 either side of the event
          if(type == "pickup"){n_pickups[cell+1] = n_pickups[cell+1] + 1}
        }
      }
    }
  }
  # End intervals
  for(cell in (which(cells_with_scooters)-1)){
    cells_with_scooters[cell+1] = FALSE
    cell_status = cell_status %>%
      add_row(cell, n_pickups = n_pickups[cell+1], nzi_start = nzi_start[cell+1], nzi_end = t_end)
    n_pickups[cell+1] = 0
    nzi_start[cell+1] = 0
  }
  list(demands = cell_status %>% mutate(month = j), cell_populations = cell_populations) #return
}

cell_populations = demands[[2]]
demands_table = demands[[1]]

for(k in 3:length(demands)){
  if(k %% 2 == 0){
    cell_populations = rbind(cell_populations, demands[[k]])
  } else {
    demands_table = rbind(demands_table, demands[[k]])
  }
}

demands_table = demands_table %>% drop_na(cell)
cell_populations = cell_populations %>% drop_na(cell)
```

```{r}
cell_populations = cell_populations %>% filter(cell >= 0) %>% group_by(cell, time_window) %>% mutate(i = row_number()) %>% ungroup()
# 
# cell_populations %>% ungroup() %>% mutate(i_overall = row_number()) %>% ggplot(aes(x = i_overall, y = n_in_cell, group = cell)) + geom_line() + facet_wrap(~time_window)
# cell_populations %>% ungroup() %>% mutate(i_overall = row_number()) %>% ggplot(aes(x = i, y = n_in_cell, group = cell)) + geom_line() + facet_wrap(~time_window)

cell_populations = cell_populations %>% filter(cell >= 0) %>% group_by(cell) %>% mutate(i = row_number()) %>% ungroup()

cell_populations %>% ungroup() %>% mutate(i_overall = row_number()) %>% ggplot(aes(x = i_overall, y = n_in_cell, group = cell)) + geom_line()
cell_populations %>% ungroup() %>% mutate(i_overall = row_number()) %>% ggplot(aes(x = i, y = n_in_cell, group = cell)) + geom_line()
```

## Estimate demand
```{r}
demands_table = demands_table %>%
  mutate(demand = n_pickups/(nzi_end-nzi_start))

```

## Filter out inaccurate estimates
```{r}
demands_table %>% 
  group_by(cell) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  mutate(lat_ind = ((cell + 1) %% n), lon_ind = ((cell) %/% n) + 1,
         #demand = case_when(
         #   demand > 0.0005 ~ 0.0005,
         #   TRUE           ~ demand
         #)
  ) %>%
  ungroup() %>%
  complete(lat_ind = 1:50, lon_ind = 1:50, fill = list(demand = 0)) %>%
  mutate(demand = demand*3600) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = demand)) + 
  geom_raster() + labs(fill = "Demand (hires/hr)", x = "", y = "") + 
  #scale_fill_gradient2(low = "#173603", mid = "#6eff0d", high = "#ffffff", midpoint = 1, na.value = "#290e00") + 
  coord_fixed(ratio = 1) + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank()
  )# + facet_wrap(~month)

# Save for Stoneham
ggsave("./demand_est_ride.png", width = 5, height = 4) 
```

# Overall demand estimates by month
```{r}
demands_table %>% 
  group_by(cell, month) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  mutate(lat_ind = ((cell + 1) %% n), lon_ind = ((cell) %/% n) + 1) %>%
  group_by(month) %>%
  complete(lat_ind = 1:50, lon_ind = 1:50, fill = list(demand = 0)) %>%
  summarise(demand = sum(demand)) %>%
  mutate(month_start = times[month], demand_hr = demand*3600)
```

```{r}
demands_table %>% 
  group_by(cell, month) %>%
  summarise(n_hires = sum(n_pickups), nonzero_time = sum(nzi_end-nzi_start)/1000, demand = weighted.mean(demand, nzi_end-nzi_start)) %>%
  drop_na() %>%
  filter(n_hires > pickup_threshold, nonzero_time > time_threshold) %>%
  mutate(lat_ind = ((cell + 1) %% n), lon_ind = ((cell) %/% n) + 1) %>%
  group_by(month) %>%
  complete(lat_ind = 1:50, lon_ind = 1:50, fill = list(demand = 0)) %>%
  summarise(demand = sum(demand)) %>%
  mutate(month_start = times[month], demand_hr = demand*3600) %>%
  ggplot(aes(x = month_start, y = demand_hr)) + geom_col()
```
