---
title: "Simulated Annealing"
output: html_document
---

```{r setup, include=FALSE}
# LAST UPDATED: 2021-01-21

knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "../")
set.seed(2021)

pacman::p_load(tidyverse,
               lubridate,
               sp,
               plotly,
               foreach,
               skimr,
               optimization)

cl = doParallel::registerDoParallel(16)

options("lubridate.week.start"=1) #set first day of week to Monday instead of Sunday 
```

# Setup 

## Parameters
```{r}
n = 50
n_simulations = 16 #number of simulations to run
flat_fee = 1 
fee_per_sec = 0.38/60
n_scooters = 1300 #number of scooters in the system. Requires sum(initial_deployment) == n_scooters, otherwise returns NA. Needs to be > 1
n_steps = 1e5 #maximum number of steps in the simulation
t_final = 2*3600 #final time after o, in seconds
dt = 7200 #time step for output
variation_amount = 2 #scale for variation; on average, this many scooters move this many cells in each axis
```
## Load data

First, the binned data is loaded:
```{r main_data}
tz = "Australia/Adelaide"
trips_data = read.csv("../Likelihood Estimation/trips_binned_beam.csv", row.names = 1)
status_data = read.csv("../Likelihood Estimation/status_binned_beam.csv", row.names = 1)
```

Time descriptors:
```{r} 
timeDesc = expand_grid(time = 0:23, day = 1:7) %>%
  mutate(desc = paste(
    case_when(
      time %in% 0:5 ~ "EarlyMorning",
      time %in% 6:11 ~ "Morning",
      time %in% 12:14 ~ "EarlyAfternoon",
      time %in% 15:18 ~ "Afternoon",
      time %in% 19:23 ~ "Evening",
      TRUE ~ "NA"
    ), case_when(
      day %in% c(6,7) ~ "Weekend",
      day %in% 1:5 ~ "Weekday",
      TRUE ~ "NA"
    ),
    sep = ""
  ))

```

Demand estimates:
```{r demand_multi-time_data}
filenames = list.files(path = "../Likelihood Estimation/Demands", pattern = "beam")
filenames = filenames[!str_detect(filenames, "redists")] #not redistributions
(nTimePairs = length(filenames))
demands = array(dim = c(n, n, length(unique(timeDesc$desc))))
for(k in 1:nTimePairs){
  demands[ , ,k]  = read.csv(paste("../Likelihood Estimation/Demands/", filenames[k], sep = ""), header = FALSE) %>% as.matrix()
}
dimnames(demands)[[3]] <- gsub(".csv|_|beam|redists", "", filenames)
```

The trip time distribution parameters are loaded:
```{r triptime_names}
trip_time_params = read.csv("../Simulation/parameters_lognorm_beam.csv", row.names = 1) %>%
  as_tibble() %>%
  complete(pickup_cell_num = 0:(n^2-1), dropoff_cell_num = 0:(n^2-1), fill = list("mu" = 0, "sigma" = 0, "n_points" = 0))
```

Get trip probabilities:
```{r}
transProbs = trips_data %>% 
  group_by(pickup_cell_num) %>%
  mutate(n_pickups = n()) %>%
  group_by(dropoff_cell_num, .add = TRUE) %>%
  summarise(p = n()/mean(n_pickups))

redist_dest_probs = transProbs # TEMPORARY, NEED TO CHANGE
```

Get battery usage:
```{r}
battery_df = read.csv("../Data Analysis/battery_usage_beam.csv", row.names = 1) %>%
  as_tibble()
```


##Functions:
```{r m2_functions}
getHireTime = function(cell, demands, t, time_breaks){ #can take a vector of cells (1-n^2)
  time_desc = names(time_breaks)[min(which(time_breaks <= t))]
  l = demands[1+((cell-1)%%n), floor(1+(cell-1)/n), time_desc]
  if(l > 0){
    out = rexp(1, l)
    if(out > time_breaks[min(which(time_breaks > t))]){
      out = getHireTime(cell, demands, time_breaks[min(which(time_breaks > t))], time_breaks)
      return(out)
    } #if the hire time is after the next time break
  } else {
    out = Inf
  }
  
  return(t + out)
}

getTripTime = function(cell, dest, trip_time_params){ #cell, dest are cell indices from 1 to n^2
  rates = trip_time_params %>% filter(pickup_cell_num == cell & dropoff_cell_num == cell)
  trip_duration = exp(rnorm(1, mean = rates$mu[1], sd = rates$sigma[1]))
  return(trip_duration)
}

getNextDest = function(cell, transProbs){
  out = cell
  for(k in 1:length(cell)){
    temp = transProbs[transProbs$pickup_cell_num == cell,]
    if(nrow(temp) == 0){
      out[k] = sample(unique(transProbs$dropoff_cell_num), size = 1)
    } else if(nrow(temp) == 1){
      out[k] = temp$dropoff_cell_num
    } else {
      out[k] = sample(temp$dropoff_cell_num, size = 1, prob = temp$p)
    }
    
  }
  out = replace(out, is.nan(out), sample(c(1:n^2), 1))
  return(out)
}
```

## Function to run a simulation and return the revenue over a time window
```{r}
single_sim = function(o, dt, t_final, initial_deployment, n_steps, n, n_scooters, flat_fee, fee_per_sec, status_table){
  
  # Time breaks (for non-homogeneous demand)
  time_breaks = seq(floor_date(as.POSIXct(o, tz = tz, origin = origin), unit = "day"), by = "DSTday", length.out = ceiling(t_final/86400))
  time_breaks = rep(time_breaks, each = 5) + rep(c(0, 6*3600, 12*3600, 15*3600, 19*3600), times = length(time_breaks))
  time_desc_names = paste("demand", timeDesc$desc[timeDesc$time %in% hour(time_breaks) & timeDesc$day %in% wday(time_breaks)], sep = "")
  time_breaks = as.numeric(time_breaks)
  names(time_breaks) <- time_desc_names
  # Initialisation
  
  t = rep(o, n_steps+1) #store the times
  # t_regular = seq(from = 0, to = t_final, by = dt) #regular, hourly "snapshots". 6*3600 = 6 hours
  # index_regular = 1 #current t_regular value
  # 
  # prop_in_use = rep(0, length(t_regular))
  # 
  total_revenue = 0
  # revenue_regular = rep(0, length(t_regular))
  # 
  missed_trips = 0
  # missed_trips_regular = rep(0, length(t_regular))
  # 
  trips = 0
  # trips_regular = rep(0, length(t_regular))
  
  
  # Get first hire times
  status_table$next_hire = foreach(k = 1:n^2, .combine = c, .packages = c("lubridate")) %do% {getHireTime(k, demands, o, time_breaks)}
  
  
    for(k in 1:n_steps){
    #print(k)
    t[k] = min(status_table$next_hire, flatten_dbl(status_table$inbound_scooters_next_time))#, o + t_regular[index_regular])
    if(t[k] > o + t_final){break}
    # if(t[k] == o + t_regular[index_regular]){ # if the next time is at a reporting point
    #   prop_in_use[index_regular] = sum(sapply(status_table$inbound_scooters_id, length))/n_scooters # (number of inbound scooters)/n_scooters
    #   revenue_regular[index_regular] = total_revenue
    #   missed_trips_regular[index_regular] = missed_trips
    #   trips_regular[index_regular] = trips
    #   index_regular = index_regular + 1
    #   
    #   if(index_regular > length(t_regular)){break} #end simulations
    #   
    # } else 
    if(t[k] %in% status_table$next_hire){ # if a hire occurs
      cell = status_table$cell[status_table$next_hire == t[k]]
      cell = cell[1] #only select the first scooter to be updated
      if(status_table$n_scooters[cell] == 0){ # if there are no scooters available
        missed_trips = missed_trips + 1
        status_table$next_hire[cell] = getHireTime(cell, demands, t[k], time_breaks)
      } else { # if there is a scooter available
        # Get trip details
        trip_dest = getNextDest(cell, transProbs)
        trip_length = getTripTime(cell, trip_dest, trip_time_params)
        idle_battery_use = 0 #battery_df$battery_rate[(battery_df$cell_num == cell) & (battery_df$is_idle)]
        trip_battery_use = battery_df$battery_rate[(battery_df$cell_num == trip_dest) & (!battery_df$is_idle)] * trip_length
        valid_scooters = status_table$current_scooters_id[[cell]][(status_table$current_scooters_battery[[cell]]+idle_battery_use*(t[k]-status_table$current_scooters_last_moved[[cell]]) + trip_battery_use) > 0] #if the remaining battery is greater than the battery required for the trip
        if(length(valid_scooters) > 0){
          scooter_id = sample(valid_scooters, 1) #choose a random valid scooter
          scooter = which(status_table$current_scooters_id[[cell]] == scooter_id)
            
          # Update revenue
          total_revenue = total_revenue + flat_fee + fee_per_sec*trip_length
          
          # Update number of trips
          trips = trips + 1
          
          # Update table
          status_table$inbound_scooters_id[[trip_dest]] = c(status_table$inbound_scooters_id[[trip_dest]], status_table$current_scooters_id[[cell]][scooter])
          status_table$inbound_scooters_next_time[[trip_dest]] = c(status_table$inbound_scooters_next_time[[trip_dest]], t[k] + trip_length)
          status_table$inbound_scooters_battery[[trip_dest]] = c(status_table$inbound_scooters_battery[[trip_dest]], status_table$current_scooters_battery[[cell]][scooter] + idle_battery_use*(t[k] - status_table$current_scooters_last_moved[[cell]][scooter]) + trip_battery_use)
          status_table$n_scooters[cell] = status_table$n_scooters[cell]-1
          status_table$current_scooters_id[[cell]] = status_table$current_scooters_id[[cell]][-scooter]
          status_table$current_scooters_battery[[cell]] = status_table$current_scooters_battery[[cell]][-scooter]
          status_table$current_scooters_last_moved[[cell]] = status_table$current_scooters_last_moved[[cell]][-scooter]
          status_table$next_hire[cell] = getHireTime(cell, demands, t[k], time_breaks)
        } else { #if no scooters have enough battery
          missed_trips = missed_trips + 1
          status_table$next_hire[cell] = getHireTime(cell, demands, t[k], time_breaks)
        }
      }
      
    } else if(t[k] %in% flatten_dbl(status_table$inbound_scooters_next_time)){ # if a scooter arrives
      # find scooter to update
      cell = which(sapply(sapply(status_table$inbound_scooters_next_time, "%in%", t[k]), any))
      cell = cell[1] #only select the first scooter to be updated
      scooter = which.min(status_table$inbound_scooters_next_time[[cell]])
      
      # Update table
      status_table$current_scooters_id[[cell]] = c(status_table$current_scooters_id[[cell]], status_table$inbound_scooters_id[[cell]][scooter])
      status_table$current_scooters_battery[[cell]] = c(status_table$current_scooters_battery[[cell]], status_table$inbound_scooters_battery[[cell]][scooter])
      status_table$current_scooters_last_moved[[cell]] = c(status_table$current_scooters_last_moved[[cell]], status_table$inbound_scooters_next_time[[cell]][scooter])
      status_table$inbound_scooters_id[[cell]] = status_table$inbound_scooters_id[[cell]][-scooter]
      status_table$inbound_scooters_next_time[[cell]] = status_table$inbound_scooters_next_time[[cell]][-scooter]
      status_table$inbound_scooters_battery[[cell]] = status_table$inbound_scooters_battery[[cell]][-scooter]
      status_table$n_scooters[cell] = status_table$n_scooters[cell] + 1
    } else print(paste("something went wrong, loop", k))
    }
  return(tibble(revenue = total_revenue, trips = trips, missed_trips = missed_trips))
}
```

## Multiple simulations
```{r} 
multi_sim = function(initial_deployment){
  o = min(status_data$location_timestamp)/1000 #initial time
  
  ## Non-output array initialisation
  ### List of cells
  cells = 1:n^2
  ### Initial scooter locations
  scooter_id_list = as.character(1:n_scooters)
  scooter_locations = initial_deployment #weight by all pickup locations in the trip data (temporary, for demonstration)]
  scooter_battery = runif(n_scooters)
  scooter_lastmoved = rep(o, n_scooters)
  current_scooters_id = foreach(k = 1:n^2, .combine = c) %do% {list(c(NULL, scooter_id_list[scooter_locations == k]))}
  current_scooters_battery = foreach(k = 1:n^2, .combine = c) %do% {list(c(NULL, scooter_battery[scooter_locations == k]))}
  current_scooters_last_moved = foreach(k = 1:n^2, .combine = c) %do% {list(c(NULL, scooter_lastmoved[scooter_locations == k]))}
  ### Inbound scooters + times
  inbound_scooters_id = list(NULL)
  inbound_scooters_next_time = list(NULL)
  inbound_scooters_battery = list(NULL)
  ### Next hire times
  
  
  status_table = tibble(cell = cells, current_scooters_id, current_scooters_battery, current_scooters_last_moved, inbound_scooters_id, inbound_scooters_next_time, inbound_scooters_battery) %>% 
    mutate(n_scooters = sapply(current_scooters_id, length))
  
  out = foreach(k = 1:n_simulations, .combine = rbind, .export = c("tz", "single_sim", "getHireTime", "getNextDest", "getTripTime", "demands", "timeDesc", "transProbs", "trip_time_params", "battery_df", "n", "n_simulations", "flat_fee", "fee_per_sec", "n_scooters", "n_steps", "t_final", "dt"), .packages = c("lubridate", "tidyverse", "foreach", "sp")) %dopar% {
    df = single_sim(o, dt, t_final, initial_deployment, n_steps, n, n_scooters, flat_fee, fee_per_sec, status_table)
    df
  }
  return(out)
}
```

### Test simulation

```{r}
out = multi_sim(sample(names(table(trips_data$pickup_cell_num)), size = n_scooters, replace = TRUE, prob = table(trips_data$pickup_cell_num)))
out %>% ggplot(aes(x = revenue)) + geom_histogram()
out %>% ggplot(aes(x = missed_trips)) + geom_histogram()
out %>% ggplot(aes(x = trips)) + geom_histogram()
```

# Optimization set-up

## Initial values
```{r}
d_0 = rep(0, n^2)
scooter_locations = sample(names(table(trips_data$pickup_cell_num)), size = n_scooters, replace = TRUE, prob = table(trips_data$pickup_cell_num))
for(k in 1:n^2){d_0[k] = sum(scooter_locations == as.character(k))}
```
## Functions
```{r}
get_revenue = function(initial_deployment_vec){
  # Input checking
  if(sum(initial_deployment_vec) != n_scooters){ #must have the correct number of scooters
    return(NA)
  } else if(min(initial_deployment_vec) < 0){ #must have non-negative scooter counts
    return(NA)
  } else if(sum(round(initial_deployment_vec)-initial_deployment_vec) != 0){ #must be integer scooter counts
    return(NA)
  }
  #Convert deployment format
  initial_deployment = foreach(k = 1:n^2, .combine = c) %do% {rep(as.character(k), initial_deployment_vec[k])}
  out = multi_sim(initial_deployment)
  return(mean(out$revenue))
}

get_revenue(d_0)

variation_function = function(para_0, fun_length, rf, temp = NA){
  para_1 = foreach(k = 1:n^2, .combine = c) %do% {rep(k, para_0[k])}
  latlon = cbind((1+((para_1-1)%%n)), floor(1+(para_1-1)/n))
  
  scooters_to_move = sample(1:nrow(latlon), max(1, floor(min(n_scooters, rexp(1, 1/mean(rf)))))) #random number of scooters to shift
  for(k in scooters_to_move){
    # Select distances in each direction
    single_rf = rf[para_1[k]]
    dists = rpois(2, rf) * (rbinom(2, 1, 0.5)-0.5)*2 #variation based on rf
    latlon[k,] = pmin(50, pmax(1, latlon[k,] + dists))
  }
  para_1 = 50*(latlon[,2]-1) + latlon[,1]
  para_2 = rep(0, n^2)
  for(k in 1:n^2){para_2[k] = sum(para_1 == k)}
  return(para_2)
}
```

## Optimisation
```{r, eval = FALSE}
optim_out = optim_sa(
  fun = get_revenue,
  start = d_0,
  trace = TRUE,
  lower = rep(0, n^2),
  upper = rep(n_scooters, n^2),
  control = list(
    nlimit = 50,
    vf = variation_function
  )
)
```

```{r, eval = FALSE}
optim_out

plot(optim_out)

optim_deployment = foreach(k = 1:n^2, .combine = c) %do% {rep(as.character(k), optim_out$par[k])}
optim_deployment

table(optim_out$par-d_0)
```


Plot cell populations:
```{r, eval = FALSE}
loops = optim_out$trace %>% 
  as.tibble() %>%
  pivot_longer(cols = x_1:rf_2500)

top_cells = loops %>%
  filter(str_detect(name, "x")) %>%
  group_by(name) %>%
  summarise(n = sum(value)) %>%
  slice_max(n, n = 10) %>%
  pull(name)

loops %>% 
  filter(str_detect(name, "x"), name %in% top_cells) %>%
  ggplot(aes(x = n_outer, y = value, col = name)) + geom_line()
```
## Simulated annealing function
```{r}
#fun = objective function to maximise
#vary_function = function that takes input x_k and returns a nearby x_k+1
#x = deployment
#T = temperature
#max_iter = maximum iterations


#based on wikipedia psuedocode
simulated_annealing = function(fun, vary_function, x_0, T_0, a, max_iter){
  x = matrix(nrow = max_iter+1, ncol = 3 + n^2) #store the output
  x[1,] = c(0, x_0, T_0, fun(x_0))
  
  
  x_k = x_0
  T_k = T_0
  for(k in 1:max_iter){
    T_k = a*T_k
    x_k = vary_function(x_k, n^2, variation_amount)
    
    prev_match = apply(x[,2:(n^2+1)], 1, function(x, vec) isTRUE(all.equal(x, vec)), x_k) #check if this solution occurred previously
    if(any(prev_match)){
      f_k = mean(fun(x_k), x[max(which(prev_match)), n^2+3])
    } else { #if this is a new solution
      f_k = fun(x_k)      
    } 
    
    if(f_k > x[k, n^2+3]){ #if the new function value is higher, accept
      x[k+1,] = c(k, x_k, T_k, f_k)
    } else if(runif(1) < exp((f_k-x[k, n^2+3])/T_k)){ #if we accept a lower value
      x[k+1,] = c(k, x_k, T_k, f_k)
    } else { #do not accept, repeat the previous row
      x[k+1,] = x[k,]
      x[k+1,1] = k
    }
  }
  return(x)
}


```

Test function:
```{r}
x = simulated_annealing(get_revenue, variation_function, d_0, 100, 0.95, 1e5)

colnames(x) = c("k", paste("x_", seq(1:n^2), sep = ""), "T_k", "f_k")
x = x %>% 
  as_tibble() %>%
  pivot_longer(cols = x_1:x_2500)

top_cells = x %>%
  filter(str_detect(name, "x")) %>%
  group_by(name) %>%
  summarise(n = sum(value)) %>%
  slice_max(n, n = 10) %>%
  pull(name)

x %>% 
  filter(str_detect(name, "x"), name %in% top_cells) %>%
  ggplot(aes(x = k, y = value, col = name)) + geom_line()

x %>% 
  filter(name == "x_1") %>%
  ggplot(aes(x = k, y = f_k)) + geom_line()

x %>% 
  filter(name == "x_1") %>%
  ggplot(aes(x = k, y = f_k)) + geom_line() + scale_x_log10()
```

## Map locations
```{r}
mean_count = x %>% 
  filter(k == max(k)) %>%
  mutate(cell = 1:2500) %>%
  transmute(lat_ind = 1 + ((cell-1) %% n), 
            lon_ind = 1 + ((cell-1)-lat_ind)/n,
            n_scooters = value) %>%
  pull(n_scooters) %>%
  range() %>%
  mean(na.rm = TRUE)

x %>% 
  filter(k == max(k)) %>%
  mutate(cell = 1:2500) %>%
  transmute(lat_ind = 1 + ((cell-1) %% n), 
            lon_ind = 1 + ((cell-1)-lat_ind)/n,
            n_scooters = value) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = n_scooters)) + geom_raster() + labs(fill = "Scooter count", x = "", y = "") + 
  scale_fill_gradient2(low = "#173603", mid = "#83ff30", high = "#ffffff", midpoint = mean_count, na.value = "#173603") + 
  coord_fixed(ratio = 1) + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.background = element_blank()
  )

# Save for Stoneham
ggsave("./final_dist.png", width = 5, height = 4) 
```

Compare to initial:
```{r}
tibble(value = d_0) %>%
  mutate(cell = 1:2500) %>%
  transmute(lat_ind = 1 + ((cell-1) %% n), 
            lon_ind = 1 + ((cell-1)-lat_ind)/n,
            n_scooters = value) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = n_scooters)) + geom_raster()
```

Compare to naive:
```{r}
p_naive = apply(demands, c(1,2), mean) %>% #take mean over all time/days
  as.vector()
p_naive[((0:2499)-((0:2499) %% n))/n < 10] = 0 # Get rid of outlier values

d_naive = rep(0, n^2)
for(k in 1:n_scooters){
  cell = which.max(p_naive/sum(p_naive)-d_naive/k) #find the maximum difference between the demand (p) and current scooter proportion (d/k)
  d_naive[cell] = d_naive[cell]+1
}

sum(d_naive)
d_naive

tibble(value = d_naive) %>%
  mutate(cell = 1:2500) %>%
  transmute(lat_ind = 1 + ((cell-1) %% n), 
            lon_ind = 1 + ((cell-1)-lat_ind)/n,
            n_scooters = value) %>%
  ggplot(aes(x = lon_ind, y = lat_ind, fill = n_scooters)) + geom_raster()

get_revenue(d_naive)
```

